Galactic Chemical Evolution Models Favor an Extended Type Ia Supernova Delay-Time Distribution

### 2024-03-22 src/tex/ms.tex
[showyourwork] overleaf sync

\usepackage{showyourwork}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
% \usepackage{layouts}
\usepackage{xcolor}
\usepackage{upgreek}
\let\tablenum\relax
\usepackage{siunitx}
\usepackage{xspace}

@@

\begin{deluxetable*}{Cccl}
    \tablecaption{A summary of parameters and their fiducial values for our chemical evolution models. We omit some parameters that are unchanged from \citetalias{Johnson2021-Migration}; see their Table 1 for details.\label{tab:multizone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Fiducial Value(s)} & \colhead{Section} & \colhead{Description}
    }
    \startdata
        % Multi-zone model parameters
        R_{\rm gal}     & [0, 20] kpc   & \ref{sec:multizone-results} & Galactocentric radius \\
        \delta R_{\rm gal}  & 100 pc    & \ref{sec:multizone-results} & Width of each concentric ring \\
        \Delta R_{\rm gal}  & N/A       & \ref{app:migration} & Change in orbital radius due to stellar migration \\
        p(\Delta R_{\rm gal}|\uptau,R_{\rm form}) & Equation \ref{eq:radial-migration}    & \ref{app:migration} & Probability density function of radial migration distance \\
        z                   & [-3, 3] kpc                & \ref{app:migration} & Distance from Galactic midplane at present day \\
        p(z|\uptau,R_{\rm final}) & Equation \ref{eq:sech-pdf}            & \ref{app:migration} & Probability density function of Galactic midplane distance\\
        \Delta t        & 10 Myr    & \ref{sec:multizone-results} & Time-step size \\
        t_{\rm max}     & 13.2 Gyr  & \ref{sec:multizone-results} & Maximum simulation time \\
        n               & 8         & \ref{sec:multizone-results} & Number of stellar populations formed per ring per time-step \\
        R_{\rm SF}      & 15.5 kpc  & \ref{sec:multizone-results} & Maximum radius of star formation \\
        M_{g,0}   & 0         & \ref{sec:sfh}     & Initial gas mass \\
        \dot M_r    & continuous    & \ref{sec:multizone-results} & Recycling rate \citep[][Equation 2]{JohnsonWeinberg2020-Starbursts} \\
        \hline
        % DTD
        R_{\rm Ia}(t)   & Equation \ref{eq:dtd-function}    & \ref{sec:dtd-models}  & delay-time distribution of Type Ia supernovae \\
        t_D             & 40 Myr    & \ref{sec:dtd-models}  & Minimum SN Ia delay time \\
        N_{\rm Ia}/M_\star  & $2.2\times10^{-3}$ M$_\odot^{-1}$ & \ref{sec:yields}  & SNe Ia per unit mass of stars formed \citep{MaozMannucci2012-SNeIaReview} \\
        \hline
        % Nucleosynthetic yields
        y_{\rm O}^{\rm CC}  & 0.015     & \ref{sec:yields}  & CCSN yield of O    \\
        y_{\rm Fe}^{\rm CC} & 0.0012    & \ref{sec:yields}  & CCSN yield of Fe   \\
        y_{\rm O}^{\rm Ia}  & 0         & \ref{sec:yields}  & SN Ia yield of O       \\
        y_{\rm Fe}^{\rm Ia} & 0.00214   & \ref{sec:yields}  & SN Ia yield of Fe \\
        \hline
        % Star formation histories
        f_{\rm IO}(t|R_{\rm gal})   & Equation \ref{eq:insideout-sfh}   & \ref{sec:sfh} & Time-dependence of the inside-out SFR \\
        f_{\rm LB}(t|R_{\rm gal})   & Equation \ref{eq:lateburst-sfh}   & \ref{sec:sfh} & Time-dependence of the late-burst SFR \\
        \uptau_{\rm rise}             & 2 Gyr     & \ref{sec:sfh} & SFR rise timescale for inside-out and early-burst models \\
        \uptau_{\rm EB}(t)          & Equation \ref{eq:earlyburst-taustar}  & \ref{sec:sfh}   & Time-dependence of the early-burst SFE timescale \\
        f_{\rm EB}(t|R_{\rm gal})   & Equation \ref{eq:earlyburst-ifr}  & \ref{sec:sfh} & Time-dependence of the early-burst IFR \\
        f_{\rm TI}(t|R_{\rm gal})   & Equation \ref{eq:twoinfall-ifr}   & \ref{sec:sfh} & Time-dependence of the two-infall IFR \\
        \hline
        % One-zone parameters
        \uptau_\star                    & 2 Gyr & \ref{sec:onezone-results} & SFE timescale in one-zone models \\
        \eta(R_{\rm gal}=8\,{\rm kpc})  & 2.15  & \ref{sec:onezone-results} & Outflow mass-loading factor at the solar annulus \\
        \uptau_{\rm sfh}(R_{\rm gal}=8\,{\rm kpc})    & 15.1 Gyr  & \ref{sec:sfh} & SFH timescale at the solar annulus \\
    \enddata
\end{deluxetable*}
\vspace{-24pt}

@@

% \begin{deluxetable*}{lcll}
% \tablecaption{Summary of SN Ia DTDs explored in this paper.\label{tab:dtds}}
% \tablehead{
% \colhead{Model} & \colhead{Eq.} & \colhead{Parameters} & \colhead{Similar to}
% }
% \startdata
\begin{table*}
    \centering
    \caption{Summary of SN Ia delay-time distributions (DTDs) explored in this paper.}
    \label{tab:dtds}
    \begin{tabular}{lcll}
        \hline\hline
        \multicolumn{1}{c}{Model} & Eq. & \multicolumn{1}{c}{Parameters} & \multicolumn{1}{c}{Similar to} \\
        \hline
        Two-population  & \ref{eq:prompt-dtd}   & $t_{\rm max}=0.05$ Gyr, $\sigma=0.015$ Gyr, & \citet{Mannucci2006-TwoPopulations} \\
                        &                       & $\uptau_{\rm Ia}=3$ Gyr & \\
        Power-law   & \ref{eq:powerlaw-dtd} & $\alpha=-1.4$                 & \citet[][cluster]{Maoz2017-CosmicDTD}; 
                                                      \citet{Heringer2019-FieldGalaxyDTD}       \\
        Power-law   & \ref{eq:powerlaw-dtd} & $\alpha=-1.1$                 & \citet[][field]{Maoz2017-CosmicDTD}; 
                                                      \citet{Wiseman2021-DESRates}              \\
        Exponential & \ref{eq:exponential-dtd}  & $\uptau_{\rm Ia}=1.5$ Gyr    & \citet[][SD]{Greggio2005-AnalyticalRates};
                                                                      \citet{Schonrich2009-RadialMixing};       \\
                    &                           &                   & \citet{Weinberg2017-ChemicalEquilibrium}  \\
        Exponential & \ref{eq:exponential-dtd}  & $\uptau_{\rm Ia}=3$ Gyr   & --- \\
        Plateau     & \ref{eq:plateau-dtd}  & $W=0.3$ Gyr, $\alpha=-1.1$    & \citet[][CLOSE DD]{Greggio2005-AnalyticalRates} \\
        Plateau     & \ref{eq:plateau-dtd}  & $W=1$ Gyr, $\alpha=-1.1$      & \citet[][WIDE DD]{Greggio2005-AnalyticalRates} \\
        Triple-system   & \ref{eq:triple-dtd}   & $f_{\rm init}=0.05f_{\rm peak}$, $t_{\rm rise}=0.5$ Gyr, & \citet{Rajamuthukumar2023-TripleEvolution} \\
                        &                       & $W=0.5$ Gyr, $\alpha=-1.1$ & \\
        \hline
    \end{tabular}
\end{table*}
% \enddata
% \end{deluxetable*}

\paragraph{Two-population} A DTD in which $\sim50\%$ of SNe Ia belong to a ``prompt'' Gaussian component at small $t$ and the remainder form an exponential tail at large $t$:
\begin{equation}
    f_{\rm Ia}^{\rm twopop}(t) = 0.5\Big[e^{-\frac{(t-t_p)^2}{2\sigma^2}} + e^{-t/\uptau}\Big]
    \label{eq:prompt-dtd}
\end{equation}
To approximate the DTD from \citet{Mannucci2006-TwoPopulations}, we take $t_p=50$ Myr, $\sigma=15$ Myr, and $\uptau_{\rm Ia}=3$ Gyr, which results in $\sim 50\%$ of SNe Ia exploding within $t<100$ Myr. As we illustrate in Figure \ref{fig:dtds}, the two-population DTD has a shorter median delay time than most other models (except the power-law with $\alpha=-1.4$, not shown).
This formulation is slightly different than the approximation used in other GCE studies \citep[e.g.,][]{Matteucci2006-BimodalDTDConsequences,Poulhazan2018-PrecisionPollution}, where it has a more distinctly bimodal shape. We have compared the two approximations to this DTD in a one-zone model and found that they produce very similar abundance tracks and MDFs. This DTD is adopted by the Feedback In Realistic Environments \citep[FIRE;][]{Hopkins2014-FIRE-1} and FIRE-2 \citep{Hopkins2018-FIRE-2} simulations.

@@

\paragraph{Exponential} An exponentially declining DTD with timescale $\uptau_{\rm Ia}$:
\begin{equation}
    f_{\rm Ia}^{\rm exp}(t) = e^{-t/\uptau_{\rm Ia}}
    \label{eq:exponential-dtd}
\end{equation}
This model allows analytic solutions to the abundances as a function of time for some SFHs, making it a popular choice. \citet{Schonrich2009-RadialMixing} and \citet{Weinberg2017-ChemicalEquilibrium} both assume an exponential DTD with a timescale $\uptau_{\rm Ia}=1.5$ Gyr. We show in Appendix \ref{app:analytical-dtds} that a $\uptau_{\rm Ia}=1.5$ Gyr exponential is an adequate approximation for the analytical SD DTD from \citet{Greggio2005-AnalyticalRates}. 
There is less observational support for an exponential DTD. \citet{Strolger2020-ExponentialDTD}, fitting to the cosmic SFH and SFHs from field galaxies, find a range of exponential-like solutions with timescales $\sim 1.5 - 6$ Gyr.
In this paper, we investigate timescales $\uptau_{\rm Ia}=1.5$ and 3 Gyr.

@@

The inside-out and late-burst models
% The first two models are taken from \citetalias{Johnson2021-Migration} and 
are run in \vice's ``star formation mode,'' where the SFR surface density $\dot\Sigma_\star$ is prescribed along with the star formation efficiency (SFE) timescale $\uptau_\star\equiv \Sigma_g/\dot\Sigma_\star$. The remaining quantities, infall rate (IFR) surface density $\dot\Sigma_{\rm in}$ and gas surface density $\Sigma_g$, are calculated from the specified quantities. The latter two models are run in ``infall mode,'' where we specify $\dot\Sigma_{\rm in}$, $\uptau_\star$, and an initial mass of the ISM at the onset of star formation, which we assume to be zero for all models (including those run in SFR mode). The mode in which \vice models are run makes no difference as a unique solution can always be obtained if two of the four parametric forms are specified. 
% See Appendix B of \citetalias{Johnson2021-Migration} for details on the normalization of the SFH. 
The SFH is normalized such that the model predicts an accurate total stellar mass and surface density gradient (see Appendix B of \citetalias{Johnson2021-Migration}).
We present an overview of the four SFHs in Figure \ref{fig:sfhs}, and we discuss them individually here.

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/star_formation_histories.pdf}
    \caption{The surface densities of star formation $\dot \Sigma_\star$ (first row), gas infall $\dot \Sigma_{\rm in}$ (second row), and gas mass $\Sigma_g$ (third row), and the SFE timescale $\uptau_\star$ (fourth row) as functions of simulation time for our four model SFHs: inside-out (first column; see Equation \ref{eq:insideout-sfh}), late-burst (second column; see Equation \ref{eq:lateburst-sfh}), early-burst (third column; see Equations \ref{eq:earlyburst-ifr} and \ref{eq:earlyburst-taustar}), and two-infall (fourth column; see Equation \ref{eq:twoinfall-ifr}). In each panel, we plot curves for the model zones which have inner radii at 4 kpc (yellow), 6 kpc (orange), 8 kpc (red), 10 kpc (violet), 12 kpc (indigo), and 14 kpc (blue).}
    \label{fig:sfhs}
    \script{star_formation_histories.py}
\end{figure*}

\paragraph{Inside-out} As in \citetalias{Johnson2021-Migration}, this is our fiducial SFH. The dimensionless time-dependence of the SFR is given by
\begin{equation}
    f_{\rm IO}(t|R_{\rm gal}) = (1 - e^{-t/\uptau_{\rm rise}}) e^{-t/\uptau_{\rm sfh}},
    \label{eq:insideout-sfh}
\end{equation}
where we assume $\uptau_{\rm rise}=2$ Gyr for all radii. The SFH timescale $\uptau_{\rm sfh}$ varies with $R_{\rm gal}$, with $\uptau_{\rm sfh}(R_{\rm gal}=8\,\rm{kpc})\approx15$ Gyr at the solar annulus and longer timescales in the outer Galaxy. The $\uptau_{\rm sfh} - R_{\rm gal}$ relation is based on the radial gradients in stellar age in Milky Way-like spirals measured by \citet[][see Section 2.5 of \citetalias{Johnson2021-Migration} for details]{Sanchez2020-StarFormationTimescales}.

\paragraph{Late-burst} A variation on the inside-out SFH with a burst in the SFR at late times which is described by a Gaussian according to
\begin{equation}
    f_{\rm LB}(t|R_{\rm gal}) = f_{\rm IO}(t|R_{\rm gal}) \Big(1 + A_b e^{-(t-t_b)^2/2\sigma_b^2} \Big),
    \label{eq:lateburst-sfh}
\end{equation}
where $A_b$ is the dimensionless amplitude of the starburst, $t_b$ is the time of the peak of the burst, and $\sigma_b$ is the width of the Gaussian. 
Evidence for a recent star formation burst $\sim 2-3$ Gyr ago has been found in {\it Gaia} \citep{Mor2019-Starburst} and in massive WDs in the solar neighborhood \citep{Isern2019-Starburst}.
Following \citetalias{Johnson2021-Migration}, we adopt $A_b=1.5$, $t_b=11.2$ Gyr, and $\sigma_b=1$ Gyr. The determination of $\uptau_{\rm sfh}$ 
% and the normalization of the SFR as a function of $R_{\rm gal}$ 
is the same as in the inside-out case.

\paragraph{Early-burst} An extension of the model proposed by \citet{Conroy2022-ThickDisk} to explain the non-monotonic behavior of the high-$\alpha$ sequence down to ${\rm [Fe/H]}\approx-2.5$. This model features an abrupt factor $\sim20$ rise in the SFE at early times, driving an increase in the [O/Fe] abundance at the transition between the epochs of halo and thick disk formation. \citet{Stahlholdt2022-StarFormationEpochs} find evidence for a burst $\sim10$ Gyr ago which marks the beginning of a second phase of star formation. [Something about $\alpha$-bimodality from another paper]. We adopt the following formula for the time-dependence of the SFE timescale from \citet{Conroy2022-ThickDisk}:
\begin{equation}
    \frac{\uptau_{\rm EB}}{1\,\rm{Gyr}} =
    \begin{cases}
        50, & t < 2.5\,\rm{Gyr} \\
        \frac{50}{[1+3(t-2.5)]^2}, & 2.5\leq t \leq 3.7\,\rm{Gyr} \\
        2.36, & t > 3.7\,\rm{Gyr}.
    \end{cases}
    \label{eq:earlyburst-taustar}
\end{equation}
While \citet{Conroy2022-ThickDisk} use a constant infall rate in their one-zone model, we adopt a radially-dependent infall rate which declines exponentially with time:
\begin{equation}
    f_{\rm EB}(t|R_{\rm gal}) = e^{-t/\uptau_{\rm sfh}},
    \label{eq:earlyburst-ifr}
\end{equation}
where $\uptau_{\rm sfh}$ is the same as in the inside-out case.
To calculate $\dot \Sigma_\star$ from the above quantities, we modify the fiducial star formation law adopted from \citetalias{Johnson2021-Migration}, substituting $\uptau_{\rm EB}$ for the SFE timescale of molecular gas:
\begin{equation}
    \dot \Sigma_\star = 
    \begin{cases}
        \Sigma_g \uptau_{\rm EB}^{-1}, & \Sigma_g \geq \Sigma_{g,2} \\
        \Sigma_g \uptau_{\rm EB}^{-1} \Big(\frac{\Sigma_g}{\Sigma_{g,2}}\Big)^{2.6}, & \Sigma_{g,1} \leq \Sigma_g \leq \Sigma_{g,2} \\
        \Sigma_g \uptau_{\rm EB}^{-1} \Big(\frac{\Sigma_{g,1}}{\Sigma_{g,2}}\Big)^{2.6} \Big(\frac{\Sigma_g}{\Sigma_{g,1}}\Big)^{0.7}, & \Sigma_g \leq \Sigma_{g,1},
    \end{cases}
    \label{eq:earlyburst-sfr}
\end{equation}
with $\Sigma_{g,1}=5\times 10^6\,{\rm M}_\odot\,{\rm kpc}^{-2}$ and $\Sigma_{g,2}=2\times 10^7\,{\rm M}_\odot\,{\rm kpc}^{-2}$.

>>>>

\usepackage{showyourwork}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
% \usepackage{layouts}
\usepackage{xcolor}
% \usepackage{upgreek}
\let\tablenum\relax
\usepackage{siunitx}
\usepackage{xspace}

@@

\begin{deluxetable*}{Cccl}
    \tablecaption{A summary of parameters and their fiducial values for our chemical evolution models. We omit some parameters that are unchanged from \citetalias{Johnson2021-Migration}; see their Table 1 for details.\label{tab:multizone-parameters}}
    \tablehead{
        \colhead{Quantity} & \colhead{Fiducial Value(s)} & \colhead{Section} & \colhead{Description}
    }
    \startdata
        % Multi-zone model parameters
        R_{\rm gal}     & [0, 20] kpc   & \ref{sec:multizone-results} & Galactocentric radius \\
        \delta R_{\rm gal}  & 100 pc    & \ref{sec:multizone-results} & Width of each concentric ring \\
        \Delta R_{\rm gal}  & N/A       & \ref{app:migration} & Change in orbital radius due to stellar migration \\
        p(\Delta R_{\rm gal}|\tau,R_{\rm form}) & Equation \ref{eq:radial-migration}    & \ref{app:migration} & Probability density function of radial migration distance \\
        z                   & [-3, 3] kpc                & \ref{app:migration} & Distance from Galactic midplane at present day \\
        p(z|\tau,R_{\rm final}) & Equation \ref{eq:sech-pdf}            & \ref{app:migration} & Probability density function of Galactic midplane distance\\
        \Delta t        & 10 Myr    & \ref{sec:multizone-results} & Time-step size \\
        t_{\rm max}     & 13.2 Gyr  & \ref{sec:multizone-results} & Maximum simulation time \\
        n               & 8         & \ref{sec:multizone-results} & Number of stellar populations formed per ring per time-step \\
        R_{\rm SF}      & 15.5 kpc  & \ref{sec:multizone-results} & Maximum radius of star formation \\
        M_{g,0}   & 0         & \ref{sec:sfh}     & Initial gas mass \\
        \dot M_r    & continuous    & \ref{sec:multizone-results} & Recycling rate \citep[][Equation 2]{JohnsonWeinberg2020-Starbursts} \\
        \hline
        % DTD
        R_{\rm Ia}(t)   & Equation \ref{eq:dtd-function}    & \ref{sec:dtd-models}  & delay-time distribution of Type Ia supernovae \\
        t_D             & 40 Myr    & \ref{sec:dtd-models}  & Minimum SN Ia delay time \\
        N_{\rm Ia}/M_\star  & $2.2\times10^{-3}$ M$_\odot^{-1}$ & \ref{sec:yields}  & SNe Ia per unit mass of stars formed \citep{MaozMannucci2012-SNeIaReview} \\
        \hline
        % Nucleosynthetic yields
        y_{\rm O}^{\rm CC}  & 0.015     & \ref{sec:yields}  & CCSN yield of O    \\
        y_{\rm Fe}^{\rm CC} & 0.0012    & \ref{sec:yields}  & CCSN yield of Fe   \\
        y_{\rm O}^{\rm Ia}  & 0         & \ref{sec:yields}  & SN Ia yield of O       \\
        y_{\rm Fe}^{\rm Ia} & 0.00214   & \ref{sec:yields}  & SN Ia yield of Fe \\
        \hline
        % Star formation histories
        f_{\rm IO}(t|R_{\rm gal})   & Equation \ref{eq:insideout-sfh}   & \ref{sec:sfh} & Time-dependence of the inside-out SFR \\
        f_{\rm LB}(t|R_{\rm gal})   & Equation \ref{eq:lateburst-sfh}   & \ref{sec:sfh} & Time-dependence of the late-burst SFR \\
        \tau_{\rm rise}             & 2 Gyr     & \ref{sec:sfh} & SFR rise timescale for inside-out and early-burst models \\
        \tau_{\rm EB}(t)          & Equation \ref{eq:earlyburst-taustar}  & \ref{sec:sfh}   & Time-dependence of the early-burst SFE timescale \\
        f_{\rm EB}(t|R_{\rm gal})   & Equation \ref{eq:earlyburst-ifr}  & \ref{sec:sfh} & Time-dependence of the early-burst IFR \\
        f_{\rm TI}(t|R_{\rm gal})   & Equation \ref{eq:twoinfall-ifr}   & \ref{sec:sfh} & Time-dependence of the two-infall IFR \\
        \hline
        % One-zone parameters
        \tau_\star                    & 2 Gyr & \ref{sec:onezone-results} & SFE timescale in one-zone models \\
        \eta(R_{\rm gal}=8\,{\rm kpc})  & 2.15  & \ref{sec:onezone-results} & Outflow mass-loading factor at the solar annulus \\
        \tau_{\rm sfh}(R_{\rm gal}=8\,{\rm kpc})    & 15.1 Gyr  & \ref{sec:sfh} & SFH timescale at the solar annulus \\
    \enddata
\end{deluxetable*}
\vspace{-24pt}

@@

% \begin{deluxetable*}{lcll}
% \tablecaption{Summary of SN Ia DTDs explored in this paper.\label{tab:dtds}}
% \tablehead{
% \colhead{Model} & \colhead{Eq.} & \colhead{Parameters} & \colhead{Similar to}
% }
% \startdata
\begin{table*}
    \centering
    \caption{Summary of SN Ia delay-time distributions (DTDs) explored in this paper.}
    \label{tab:dtds}
    \begin{tabular}{lcll}
        \hline\hline
        \multicolumn{1}{c}{Model} & Eq. & \multicolumn{1}{c}{Parameters} & \multicolumn{1}{c}{Similar to} \\
        \hline
        Two-population  & \ref{eq:prompt-dtd}   & $t_{\rm max}=0.05$ Gyr, $\sigma=0.015$ Gyr, & \citet{Mannucci2006-TwoPopulations} \\
                        &                       & $\tau=3$ Gyr & \\
        Power-law   & \ref{eq:powerlaw-dtd} & $\alpha=-1.4$                 & \citet[][cluster]{Maoz2017-CosmicDTD}; 
                                                      \citet{Heringer2019-FieldGalaxyDTD}       \\
        Power-law   & \ref{eq:powerlaw-dtd} & $\alpha=-1.1$                 & \citet[][field]{Maoz2017-CosmicDTD}; 
                                                      \citet{Wiseman2021-DESRates}              \\
        Exponential & \ref{eq:exponential-dtd}  & $\tau=1.5$ Gyr    & \citet[][SD]{Greggio2005-AnalyticalRates};
                                                                      \citet{Schonrich2009-RadialMixing};       \\
                    &                           &                   & \citet{Weinberg2017-ChemicalEquilibrium}  \\
        Exponential & \ref{eq:exponential-dtd}  & $\tau=3$ Gyr   & --- \\
        Plateau     & \ref{eq:plateau-dtd}  & $W=0.3$ Gyr, $\alpha=-1.1$    & \citet[][CLOSE DD]{Greggio2005-AnalyticalRates} \\
        Plateau     & \ref{eq:plateau-dtd}  & $W=1$ Gyr, $\alpha=-1.1$      & \citet[][WIDE DD]{Greggio2005-AnalyticalRates} \\
        Triple-system   & \ref{eq:triple-dtd}   & $f_{\rm init}=0.05f_{\rm peak}$, $t_{\rm rise}=0.5$ Gyr, & \citet{Rajamuthukumar2023-TripleEvolution} \\
                        &                       & $W=0.5$ Gyr, $\alpha=-1.1$ & \\
        \hline
    \end{tabular}
\end{table*}
% \enddata
% \end{deluxetable*}

\paragraph{Two-population} A DTD in which $\sim50\%$ of SNe Ia belong to a ``prompt'' Gaussian component at small $t$ and the remainder form an exponential tail at large $t$:
\begin{equation}
    f_{\rm Ia}^{\rm twopop}(t) = 0.5\Big[e^{-\frac{(t-t_p)^2}{2\sigma^2}} + e^{-t/\tau}\Big]
    \label{eq:prompt-dtd}
\end{equation}
To approximate the DTD from \citet{Mannucci2006-TwoPopulations}, we take $t_p=50$ Myr, $\sigma=15$ Myr, and $\tau=3$ Gyr, which results in $\sim 50\%$ of SNe Ia exploding within $t<100$ Myr. As we illustrate in Figure \ref{fig:dtds}, the two-population DTD has a shorter median delay time than most other models (except the power-law with $\alpha=-1.4$, not shown).
This formulation is slightly different than the approximation used in other GCE studies \citep[e.g.,][]{Matteucci2006-BimodalDTDConsequences,Poulhazan2018-PrecisionPollution}, where it has a more distinctly bimodal shape. We have compared the two approximations to this DTD in a one-zone model and found that they produce very similar abundance tracks and MDFs. This DTD is adopted by the Feedback In Realistic Environments \citep[FIRE;][]{Hopkins2014-FIRE-1} and FIRE-2 \citep{Hopkins2018-FIRE-2} simulations.

@@

\paragraph{Exponential} An exponentially declining DTD with timescale $\tau$:
\begin{equation}
    f_{\rm Ia}^{\rm exp}(t) = e^{-t/\tau}
    \label{eq:exponential-dtd}
\end{equation}
This model allows analytic solutions to the abundances as a function of time for some SFHs, making it a popular choice. \citet{Schonrich2009-RadialMixing} and \citet{Weinberg2017-ChemicalEquilibrium} both assume an exponential DTD with a timescale $\tau=1.5$ Gyr. We show in Appendix \ref{app:analytical-dtds} that a $\tau=1.5$ Gyr exponential is an adequate approximation for the analytical SD DTD from \citet{Greggio2005-AnalyticalRates}. 
There is less observational support for an exponential DTD. \citet{Strolger2020-ExponentialDTD}, fitting to the cosmic SFH and SFHs from field galaxies, find a range of exponential-like solutions with timescales $\sim 1.5 - 6$ Gyr.
In this paper, we investigate timescales $\tau=1.5$ and 3 Gyr.

@@

The inside-out and late-burst models
% The first two models are taken from \citetalias{Johnson2021-Migration} and 
are run in \vice's ``star formation mode,'' where the SFR surface density $\dot\Sigma_\star$ is prescribed along with the star formation efficiency (SFE) timescale $\tau_\star\equiv \Sigma_g/\dot\Sigma_\star$. The remaining quantities, infall rate (IFR) surface density $\dot\Sigma_{\rm in}$ and gas surface density $\Sigma_g$, are calculated from the specified quantities. The latter two models are run in ``infall mode,'' where we specify $\dot\Sigma_{\rm in}$, $\tau_\star$, and an initial mass of the ISM at the onset of star formation, which we assume to be zero for all models (including those run in SFR mode). The mode in which \vice models are run makes no difference as a unique solution can always be obtained if two of the four parametric forms are specified. 
% See Appendix B of \citetalias{Johnson2021-Migration} for details on the normalization of the SFH. 
The SFH is normalized such that the model predicts an accurate total stellar mass and surface density gradient (see Appendix B of \citetalias{Johnson2021-Migration}).
We present an overview of the four SFHs in Figure \ref{fig:sfhs}, and we discuss them individually here.

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/star_formation_histories.pdf}
    \caption{The surface densities of star formation $\dot \Sigma_\star$ (first row), gas infall $\dot \Sigma_{\rm in}$ (second row), and gas mass $\Sigma_g$ (third row), and the SFE timescale $\tau_\star$ (fourth row) as functions of simulation time for our four model SFHs: inside-out (first column; see Equation \ref{eq:insideout-sfh}), late-burst (second column; see Equation \ref{eq:lateburst-sfh}), early-burst (third column; see Equations \ref{eq:earlyburst-ifr} and \ref{eq:earlyburst-taustar}), and two-infall (fourth column; see Equation \ref{eq:twoinfall-ifr}). In each panel, we plot curves for the model zones which have inner radii at 4 kpc (yellow), 6 kpc (orange), 8 kpc (red), 10 kpc (violet), 12 kpc (indigo), and 14 kpc (blue).}
    \label{fig:sfhs}
    \script{star_formation_histories.py}
\end{figure*}

\paragraph{Inside-out} As in \citetalias{Johnson2021-Migration}, this is our fiducial SFH. The dimensionless time-dependence of the SFR is given by
\begin{equation}
    f_{\rm IO}(t|R_{\rm gal}) = (1 - e^{-t/\tau_{\rm rise}}) e^{-t/\tau_{\rm sfh}},
    \label{eq:insideout-sfh}
\end{equation}
where we assume $\tau_{\rm rise}=2$ Gyr for all radii. The SFH timescale $\tau_{\rm sfh}$ varies with $R_{\rm gal}$, with $\tau_{\rm sfh}(R_{\rm gal}=8\,\rm{kpc})\approx15$ Gyr at the solar annulus and longer timescales in the outer Galaxy. The $\tau_{\rm sfh} - R_{\rm gal}$ relation is based on the radial gradients in stellar age in Milky Way-like spirals measured by \citet[][see Section 2.5 of \citetalias{Johnson2021-Migration} for details]{Sanchez2020-StarFormationTimescales}.

\paragraph{Late-burst} A variation on the inside-out SFH with a burst in the SFR at late times which is described by a Gaussian according to
\begin{equation}
    f_{\rm LB}(t|R_{\rm gal}) = f_{\rm IO}(t|R_{\rm gal}) \Big(1 + A_b e^{-(t-t_b)^2/2\sigma_b^2} \Big),
    \label{eq:lateburst-sfh}
\end{equation}
where $A_b$ is the dimensionless amplitude of the starburst, $t_b$ is the time of the peak of the burst, and $\sigma_b$ is the width of the Gaussian. 
Evidence for a recent star formation burst $\sim 2-3$ Gyr ago has been found in {\it Gaia} \citep{Mor2019-Starburst} and in massive WDs in the solar neighborhood \citep{Isern2019-Starburst}.
Following \citetalias{Johnson2021-Migration}, we adopt $A_b=1.5$, $t_b=11.2$ Gyr, and $\sigma_b=1$ Gyr. The determination of $\tau_{\rm sfh}$ 
% and the normalization of the SFR as a function of $R_{\rm gal}$ 
is the same as in the inside-out case.

\paragraph{Early-burst} An extension of the model proposed by \citet{Conroy2022-ThickDisk} to explain the non-monotonic behavior of the high-$\alpha$ sequence down to ${\rm [Fe/H]}\approx-2.5$. This model features an abrupt factor $\sim20$ rise in the SFE at early times, driving an increase in the [O/Fe] abundance at the transition between the epochs of halo and thick disk formation. \citet{Stahlholdt2022-StarFormationEpochs} find evidence for a burst $\sim10$ Gyr ago which marks the beginning of a second phase of star formation. [Something about $\alpha$-bimodality from another paper]. We adopt the following formula for the time-dependence of the SFE timescale from \citet{Conroy2022-ThickDisk}:
\begin{equation}
    \frac{\tau_{\rm EB}}{1\,\rm{Gyr}} =
    \begin{cases}
        50, & t < 2.5\,\rm{Gyr} \\
        \frac{50}{[1+3(t-2.5)]^2}, & 2.5\leq t \leq 3.7\,\rm{Gyr} \\
        2.36, & t > 3.7\,\rm{Gyr}.
    \end{cases}
    \label{eq:earlyburst-taustar}
\end{equation}
While \citet{Conroy2022-ThickDisk} use a constant infall rate in their one-zone model, we adopt a radially-dependent infall rate which declines exponentially with time:
\begin{equation}
    f_{\rm EB}(t|R_{\rm gal}) = e^{-t/\tau_{\rm sfh}},
    \label{eq:earlyburst-ifr}
\end{equation}
where $\tau_{\rm sfh}$ is the same as in the inside-out case.
To calculate $\dot \Sigma_\star$ from the above quantities, we modify the fiducial star formation law adopted from \citetalias{Johnson2021-Migration}, substituting $\tau_{\rm EB}$ for the SFE timescale of molecular gas:
\begin{equation}
    \dot \Sigma_\star = 
    \begin{cases}
        \Sigma_g \tau_{\rm EB}^{-1}, & \Sigma_g \geq \Sigma_{g,2} \\
        \Sigma_g \tau_{\rm EB}^{-1} \Big(\frac{\Sigma_g}{\Sigma_{g,2}}\Big)^{2.6}, & \Sigma_{g,1} \leq \Sigma_g \leq \Sigma_{g,2} \\
        \Sigma_g \tau_{\rm EB}^{-1} \Big(\frac{\Sigma_{g,1}}{\Sigma_{g,2}}\Big)^{2.6} \Big(\frac{\Sigma_g}{\Sigma_{g,1}}\Big)^{0.7}, & \Sigma_g \leq \Sigma_{g,1},
    \end{cases}
    \label{eq:earlyburst-sfr}
\end{equation}
with $\Sigma_{g,1}=5\times 10^6\,{\rm M}_\odot\,{\rm kpc}^{-2}$ and $\Sigma_{g,2}=2\times 10^7\,{\rm M}_\odot\,{\rm kpc}^{-2}$.

\paragraph{Two-infall} First proposed by \citet{Chiappini1997-TwoInfall}, this model parameterizes the infall rate as two successive, exponentially declining bursts to explain the origin of the high- and low-$\alpha$ disk populations:
\begin{equation}
    \label{eq:twoinfall-ifr}
    f_{\rm TI}(t|R_{\rm gal}) = N_1(R_{\rm gal}) e^{-t/\tau_1} + N_2(R_{\rm gal}) e^{-(t-t_{\rm on})/\tau_2},
\end{equation}
% In this model, the first infall produces the thick disk and the second infall produces the thin disk. 
where $N_1$ and $N_2$ are the normalizations of the first and second infall, respectively, and $t_{\rm on}$ is the onset time of the second infall.
The normalization ratio $N_2/N_1$ is calculated so that the thick-to-thin-disk surface density ratio $f_\Sigma(R)=\Sigma_2(R)/\Sigma_1(R)$ is given by
\begin{equation}
    f_\Sigma(R) = f_\Sigma(0) e^{R(1/R_2 - 1/R_1)}.
\end{equation}
Following \citet{BlandHawthornGerhard2016-MilkyWayReview}, we adopt values for the thick disk scale radius $R_1=2.0$ kpc, thin disk scale radius $R_2=2.5$ kpc, and $f_\Sigma(0)=0.27$.
We note that most previous studies which use the two-infall model \citep[e.g.,][]{Spitoni2021-TwoInfall} do not consider gas outflows and instead adjust the nucleosynthetic yields to reproduce the solar abundance. We adopt radially-dependent outflows as in \citetalias{Johnson2021-Migration} (see their Section 2.4 for details) for all our SFHs, including two-infall. We discuss the implications of this difference in Section \ref{sec:two-infall-discussion}.
